---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 10,
  fig.retina = 3
)
```

# ggsegHistorical

<!-- badges: start -->
[![R-CMD-check](https://github.com/ggsegverse/ggsegHistorical/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ggsegverse/ggsegHistorical/actions/workflows/R-CMD-check.yaml)
[![r-universe](https://ggseg.r-universe.dev/badges/ggsegHistorical)](https://ggseg.r-universe.dev/ggsegHistorical)
<!-- badges: end -->

This package provides six historical cortical brain atlases digitally
reconstructed by
[Pijnenburg et al. (2021)](https://doi.org/10.1016/j.neuroimage.2021.118274)
from the [Dutch Connectome Lab](http://www.dutchconnectomelab.nl/).

## Installation

We recommend installing the ggseg-atlases through the ggseg
[r-universe](https://ggseg.r-universe.dev/ui#builds):

```{r}
#| label: install-r-universe
#| eval: false
options(repos = c(
  ggseg = "https://ggseg.r-universe.dev",
  CRAN = "https://cloud.r-project.org"
))

install.packages("ggsegHistorical")
```

You can install this package from [GitHub](https://github.com/) with:

```r
# install.packages("pak")
pak::pak("ggsegverse/ggsegHistorical")
```

## Brodmann (1909)

```{r}
#| label: brodmann
library(ggseg)
library(ggsegHistorical)

plot(brodmann())
```

## Campbell (1905)

```{r}
#| label: campbell
plot(campbell())
```

## Economo & Koskinas (1925)

```{r}
#| label: economo
plot(economo())
```

## Flechsig (1920)

```{r}
#| label: flechsig
plot(flechsig())
```

## Kleist (1934)

```{r}
#| label: kleist
plot(kleist())
```

## Smith (1907)

```{r}
#| label: smith
plot(smith())
```

## Data source

Pijnenburg R, Scholtens LH, Mantini D, & van den Heuvel MP (2021).
Digitally reconstructed cortical brain maps of the pioneers Brodmann, Campbell,
Economo and Koskinas, Flechsig, Kleist, and Smith. *NeuroImage*, 239, 118274.
