---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 10,
  fig.retina = 3
)
```

# ggsegHistorical

<!-- badges: start -->
[![R-CMD-check](https://github.com/ggsegverse/ggsegHistorical/workflows/R-CMD-check/badge.svg)](https://github.com/ggsegverse/ggsegHistorical/actions)
<!-- badges: end -->

This package provides six historical cortical brain atlases digitally
reconstructed by
[Pijnenburg et al. (2021)](https://doi.org/10.1016/j.neuroimage.2021.118274)
from the [Dutch Connectome Lab](http://www.dutchconnectomelab.nl/), formatted
for use with
[ggseg](https://ggseg.github.io/ggseg/) and
[ggseg3d](https://ggseg.github.io/ggseg3d/).

| Atlas | Function | Year | Regions/hemi | Type |
|-------|----------|------|--------------|------|
| Brodmann | `brodmann()` | 1909 | 39 | Cytoarchitectonic |
| Campbell | `campbell()` | 1905 | 17 | Histological |
| Economo & Koskinas | `economo()` | 1925 | 15 | Cytoarchitectonic |
| Flechsig | `flechsig()` | 1920 | 46 | Myelogenetic |
| Kleist | `kleist()` | 1934 | 49 | Functional |
| Smith | `smith()` | 1907 | 14 | Anatomical |

To learn how to use these atlases, please look at the documentation for
[ggseg](https://ggseg.github.io/ggseg/) and
[ggseg3d](https://ggseg.github.io/ggseg3d/).

## Installation

You can install ggsegHistorical from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("ggseg/ggsegHistorical")
```

## Atlases

```{r}
#| label: brodmann
#| fig-height: 4
library(ggseg)
library(ggsegHistorical)
library(ggplot2)

ggplot() +
  geom_brain(
    atlas = brodmann(),
    mapping = aes(fill = label),
    position = position_brain(hemi ~ view),
    show.legend = FALSE
  ) +
  ggtitle("Brodmann (1909)") +
  theme_void()
```

```{r}
#| label: campbell
#| fig-height: 4
ggplot() +
  geom_brain(
    atlas = campbell(),
    mapping = aes(fill = label),
    position = position_brain(hemi ~ view),
    show.legend = FALSE
  ) +
  ggtitle("Campbell (1905)") +
  theme_void()
```

```{r}
#| label: economo
#| fig-height: 4
ggplot() +
  geom_brain(
    atlas = economo(),
    mapping = aes(fill = label),
    position = position_brain(hemi ~ view),
    show.legend = FALSE
  ) +
  ggtitle("Economo & Koskinas (1925)") +
  theme_void()
```

```{r}
#| label: flechsig
#| fig-height: 4
ggplot() +
  geom_brain(
    atlas = flechsig(),
    mapping = aes(fill = label),
    position = position_brain(hemi ~ view),
    show.legend = FALSE
  ) +
  ggtitle("Flechsig (1920)") +
  theme_void()
```

```{r}
#| label: kleist
#| fig-height: 4
ggplot() +
  geom_brain(
    atlas = kleist(),
    mapping = aes(fill = label),
    position = position_brain(hemi ~ view),
    show.legend = FALSE
  ) +
  ggtitle("Kleist (1934)") +
  theme_void()
```

```{r}
#| label: smith
#| fig-height: 4
ggplot() +
  geom_brain(
    atlas = smith(),
    mapping = aes(fill = label),
    position = position_brain(hemi ~ view),
    show.legend = FALSE
  ) +
  ggtitle("Smith (1907)") +
  theme_void()
```

## Citation

If you use these atlases, please cite:

> Pijnenburg R, Scholtens LH, Mantini D, de Reus MA, van den Heuvel MP (2021).
> Myelo- and cytoarchitectonic microstructural and functional human cortical
> atlases reconstructed in common MRI space. *NeuroImage*, 239, 118274.
> [doi:10.1016/j.neuroimage.2021.118274](https://doi.org/10.1016/j.neuroimage.2021.118274)

Please also cite the original historical atlas you use (see `?brodmann`,
`?campbell`, etc. for individual references).

## License

The atlas data is released under a CC BY-NC-SA 4.0 license by
[Pijnenburg et al. (2021)](https://doi.org/10.1016/j.neuroimage.2021.118274)
from the [Dutch Connectome Lab](http://www.dutchconnectomelab.nl/).
